---
title: "Activity 2 - Day 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(GGally)
```

## Day 1
### Load The Data
```{r load_data, warning=FALSE}
hfi <- readr::read_csv("https://www.openintro.org/data/csv/hfi.csv")
```

### Describe pf_score
```{r describe_pf_score, warning=FALSE}
hfi %>% 
  ggplot(aes(x=pf_score)) +
  geom_histogram() + 
  theme_bw()
```
The variable `pf_score` has a negative skew, and the distribution appears to be bimodal. Very few countries have a personal freedom score of less than 5, however, there seems to be a division between many countries with a score averaging 7 and others averaging 9.

### Comparing Killed vs Jailed
```{r compare_2_variables, warning=FALSE}
hfi %>% 
  ggplot(aes(x=pf_expression_killed, y=pf_expression_jailed)) + 
  geom_point() +
  theme_bw()
```
There is a positive correlation between `pf_expression_jailed` and `pf_expression_killed`. However, the this relationship does not appear to be linear.

### Pairwise Relationships
```{r pairwise, warning=FALSE}
hfi %>% 
  select(ends_with("_score")) %>% 
  ggpairs() + 
  theme_bw()
```

These three variables appear to have a positive linear relationship, and there appears very few data points with a `pf_score` under 4 and `ef_score` over 9.

### The Multiple Linear Regression Model
```{r mlr}
m_hr_ef <- lm(pf_score ~ hf_score + ef_score, data = hfi)
tidy(m_hr_ef)
```

y = 1.464213e-11 + 2`hf_score` - 1`ef_score`

When `hf_score` and `ef_score` are 0, then `pf_score` is 0. For each unit of `hf_score` then `pf_score` will increase by 2, and for each unit of `ef_score` then `pf_score` will decrease by 1.

### Challenge: 3D Plots
```{r 3d_plot, warning=FALSE}
# Visual Idea from https://stackoverflow.com/questions/45052188/how-to-plot-3d-scatter-diagram-using-ggplot
#library(plotly)
# plot_ly(x=hfi$pf_score, y=hfi$hf_score, z=hfi$ef_score, type="scatter3d", mode="markers", color = hfi$pf_score) 
```


